Tutorial_2

### 2024-08-07 Tutorial_2_Bayesian/Tutorial_2.Rmd
add models

```{r priors}
priors <- c(
  set_prior("normal(0, 1)", class = "b"), # for beta parameters 
  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale
  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
  set_prior("lkj(2)", class = "cor") # for correlations between RE
)
```

@@

model_full_RE <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9*trial_c + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           (1 + condition*period_9*trial_c +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_full_RE")

@@

Model_full_RE took 3 hours to run.

@@

@@

```{r check-model-full-RE}
summary(model_cubic_RE)
```

@@

```{r}

@@

```{r load-model}
model_cubic_RE <- readRDS("Tutorial_2_Bayesian/models/model_cubic_RE.rds") 
```

```{r make-plot-effects}
post <-as_draws_df(model_cubic_RE) %>% # 8000 draws x 299 variables
   select(starts_with("b_")) %>%       # 8000 x 18
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial100 = -900/1000)  
# effects for trial 1000
p1congruent <- post %>% 
  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

@@

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  labs(x = "time bin", y = "effect of incongruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

>>>>

```{r priors}
priors <- c(
  set_prior("normal(0, 1)", class = "b"), # for beta parameters when using logit link

  #set_prior("skew_normal(-0.5, .95,-4)", class = "b"), # for beta parameters when using cloglog link

  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale when using logit link;

  #set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),   # weakly regularizing prior for intercept when using cloglog link (hazard values below .5 more likely than values above .5)
  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
  set_prior("lkj(2)", class = "cor") # for correlations between RE
)

inv_cloglog <- function(c){
  1-exp(-exp(c))
}
cloglog <- function(p){
  log(-log(1-p))
}

dat <- 
  tibble(c1 = rnorm(1e5,0,1), # usa as weakly reguralizing prior for beta parameters (when using logit link)
         c2 = rt(1e5,7.61)* 1.57) %>% # use as weakly reguralizing prior for interceptparameter (when using logit 
  mutate(p1 = inv_logit_scaled(c1),
         p2 = inv_logit_scaled(c2))

plot(density(dat$c1))
plot(density(dat$c2))
plot(density(dat$p1))
plot(density(dat$p2))

# cloglog
dat2 <- tibble(c1 = rnorm(1e5,-1,1.1), # cloglog intercept
              c2 = rskew_normal(1e5, -0.5, sigma = .95, alpha = -4)) %>% # beta toward .6

  mutate(p1 = inv_cloglog(c1),
         p2 = inv_cloglog(c2))

plot(density(dat2$c1))
plot(density(dat2$c2))
plot(density(dat2$p1))
plot(density(dat2$p2))

@@

model_full_RE_newprior <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9*trial_c + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           (1 + condition*period_9*trial_c +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_full_RE_newprior")

@@

Model_full_RE took between 3 and 4 hours to run.

@@

Fit a model that relaxes proportionality assumption.

```{r}
# remove unnecessary columns before fitting a model
ptb_data <- ptb_data %>% select(-c(trial_c)) # 12840 obs. of 4 variables
head(ptb_data,n=17)
summary(ptb_data)
```

```{r cubic-random-effects-model}
plan(multicore)

@@

```{r check-model-cubic-RE}
summary(model_cubic_RE)
```

#### relax all ass

```{r cubictime-quadtrial-model}
plan(multicore)

model_cubictime_quadtrial <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9*trial_c + 
                           condition*period_9*I(trial_c^2) + 
                           condition*I(period_9^2)*trial_c + 
                           condition*I(period_9^2)*I(trial_c^2) +
                           condition*I(period_9^3) +
                      (1 + condition*period_9*trial_c +
                           condition*period_9*I(trial_c^2) +
                           condition*I(period_9^2)*trial_c +
                           condition*I(period_9^2)*I(trial_c^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_cubictime_quadtrial")

@@

```{r}
model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
model_full_RE_newprior <- readRDS("Tutorial_2_Bayesian/models/model_full_RE_newprior.rds") 
fixef(model_full_RE)
fixef(model_full_RE_newprior)
```

@@

```{r load-model}
model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
```

```{r make-plot-effects}
post <-as_draws_df(model_full_RE) %>% # 8000 draws x 299 variables
   select(starts_with("b_")) %>%       # 8000 x 18
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial100 = -900/1000)  
# effects for trial 1000
p1congruent <- post %>% 
  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

@@

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
  labs(x = "time bin", y = "effect of incongruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))
